A space-time skew-$t$ model for threshold exceedances

### 2016-05-25 LaTeX/skewt-rev.tex
changes to some of the manuscript preparing for revision
manuscript updates

% cross-referencing appendices
\usepackage{xr}
\externaldocument{supplemental-orig}

@@

Our model allows for inference and predictions using the full likelihood with computing on the order of Gaussian models for large space-time datasets.
This allows us to use Bayesian methods to impute data below the threshold as well as make predictions at unobserved locations.
In a spatial setting, the multivariate skew-$t$ distribution demonstrates asymptotic dependence between observations at all sites regardless of the distance between the sites.
In order to address this concern, we introduce a random spatial partition similar to the method used by \citet{Kim2005} for non-stationary Gaussian data.
This partitioning alleviates the asymptotic spatial dependence present in the skew-$t$ distribution for sites that are far apart.

The paper is organized as follows.
Section \ref{s:spatialskew} is a brief review of the spatial skew-$t$ process.
In Section \ref{s:spatial}, we build upon the traditional skew-$t$ process by incorporating censoring to focus on tails, partitioning to remove long-range asymptotic dependence, and extending the model to space-time data.
The computing is described in Section \ref{s:hier}.
In Section \ref{s:simstudy}, we present a simulation study that examines the predictive capabilities of this model compared Gaussian amd max-stable methods.
We then compare our method to Gaussian and max-stable methods with a data analysis of ozone measurements from 800 sites throughout the US in Section \ref{s:analysis}.
The final section provides brief discussion and direction for future research.

@@

\subsection{Skew-$t$ process} \label{s:skewt}
Let $Y(\bs)$ be the observation at spatial location $\bs$ in a spatial domain of interest $\calD \in \calR^2$.
The spatial skew-$t$ process can be written
\begin{align}
  Y(\bs) = \bX(\bs)^T \bbeta + \lambda \sigma |z| + \sigma v(\bs)
\end{align}
where $\bX(\bs)$ is a set of spatial covariates at site $\bs$, $\bbeta$ is the vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a, b)$ is random scale parameter, IG is the distribution function of an inverse gamma random variable, and $v(\bs)$ is a spatial Gaussian process with mean zero, variance one, and a positive definite correlation function.

For a finite collection of locations $\bs_1, \ldots, \bs_n$, denote the vector of observations $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^T$.
After marginalizing over both $z$ and $\sigma$,
\begin{align}
  \bY \sim \text{ST}_n(\bX \bbeta, \bOmega, \balpha, 2a), \label{eq:fullmodel}
\end{align}
that is, $\bY$ follows an $n$-dimensional skew-$t$ distribution with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$, and $\bSigma$ is the positive definite correlation matrix of $[v(\bs_1), \ldots, v(\bs_n)]$.
This process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $2a$.
Furthermore, the marginal distributions at each location also follow a univariate skew-$t$ distribution \citep{Azzalini2014}.

Although any positive definite correlation function could be used, we choose to use the stationary isotropic \Matern correlation with
\begin{align}
  \text{cor}[v(\bs_1), v(\bs_2)] = \gamma I(\bs_1 = \bs_2) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right) \label{eq:matern}
\end{align}
where $\rho$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs_1 - \bs_2 ||$.
We use this parameterization for spatial correlation because the $\gamma$ parameter permits the inclusion of a nugget effect to account for non-spatial variability due to issues like measurement error.

\subsection{Extremal dependence}\label{s:extdep}
Our interest lies in spatial dependence in the tail of the skew-$t$ process.
One measure of extremal dependence is the $\chi$ statistic \citep{Coles1999}.
For a stationary and isotropic spatial process, the $\chi$ statistic for locations $\bs$ and $\bt$ separated by distance $h = || \bs - \bt ||$ with identical marginal distributions is
\begin{align}
  \chi(h) = \lim_{c \rightarrow c^*} \Pr[Y(\bs) > c | Y(\bt) > c]
\end{align}
where $c^*$ is the upper limit of the support of $Y$.
If $\chi(h) = 0$, then observations are asymptotically independent at distance $h$.
For Gaussian processes, $\chi(h) = 0$ regardless of the distance $h$, so they are not suitable for modeling asymptotically dependent extremes.
Unlike the Gaussian process, the skew-$t$ process is asymptotically dependent (the explicit expression for $\chi(h)$ is given in \ref{a:skewt}).
However, one problem with the spatial skew-$t$ process is that $\lim_{h \rightarrow \infty} \chi(h) > 0$.
This occurs because all observations, both near and far, share the same $z$ and $\sigma$ terms.
Therefore, this long-range dependence feature of the skew-$t$ process is not ideal for spatial analysis of large geographic regions where we expect only local spatial dependence.
We propose a solution to this in Section \ref{s:part}.

@@

\subsection{Partitioning to remove long-range asymptotic dependence}\label{s:part}
The motivation for the partition is that for a large spatial domain, it may not be reasonable to assume sites that are far apart demonstrate asymptotic dependence.
Modeling different levels of asymptotic dependence was discussed by \citet{Wadsworth2012} with a hybrid spatial dependence model.
\citet{Huser2014} also allow for varying asymptotic dependence across both space and time with a partition structure represented by random discs moving across the space for a random duration with a random velocity and random radius.
We handle the problem of long-range asymptotic dependence with a random partition.
As discussed in Section \ref{s:spatialskew}, the source of long-range dependence is the shared $z$ and $\sigma$.
Therefore, to alleviate this dependence, we allow $z$ and $\sigma$ to vary by site.
The model becomes
\begin{align} \label{eq:partition}
  Y(\bs) &= \bX(\bs)^T \bbeta + \lambda \sigma(\bs) |z(\bs)| + \sigma(\bs) v(\bs).
\end{align}
To model spatial variation, consider a set of spatial knots $\bw_1, \ldots, \bw_K$ from a homogeneous Poisson process with intensity $\mu$ over spatial domain $\calD \in \calR^2$.
The knots define a random partition of $\calD$ by subregions $P_{1}, \ldots, P_{K}$ defined as
\begin{align} \label{eq:subregions}
  P_{k} = \{ \bs : k = \argmin_\ell || \bs - \bw_{\ell} || \}.
\end{align}
All $z(\bs)$ and $\sigma(\bs)$ for sites in subregion $k$ are assigned common values
\begin{align}
  z(\bs) = z_{k} \quad \text{and} \quad \sigma(\bs) = \sigma_{k} \label{eq:sitezsig}
\end{align}
and the $z_k$ and $\sigma^2_k$ are distributed as $z_k \iid N(0, 1)$ and $\sigma^2_k \iid$ IG$(a, b)$.
So, within each partition, $Y(\bs)$ follows the spatial skew-$t$ process defined in Section \ref{s:spatialskew}.
Across partitions, the $Y(\bs)$ remain correlated via the correlation function for $v(\bs)$ because $v(\bs)$ spans all partitions.

The partitioning model remove long-range dependence. Conditional on knots $\bw_1, \ldots, \bw_K$, the $\chi$ statistic for two sites $\bs$ and $\bt$ in partitions $k_s$ and $k_t$ respectively is
\begin{align}
  \chi(h) &= I(k_s = k_t) \chi_{\text{skew-}t}(h) + I(k_s \neq k_t) \chi_{\text{Gaus}}(h) \nonumber \\
         &= I(k_s = k_t) \chi_{\text{skew-}t}(h)
\end{align}
where $I(\cdot)$ is an indicator function, $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process given in (\ref{eq:chiskew-t}), $\chi_{\text{Gaus}}(h)$ is the $\chi$ statistic for a Gaussian process, and $h = ||\bs - \bt||$.
Therefore, sites in different subregions are asymptotically independent because $\chi_{\text{Gaus}}(h) = 0$ for all $h$.
Marginally, over the knots, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_s = k_t)$ is the probability that two sites separated by distance $h$ are in the same partition.
In \ref{a:proofsamepartition}, we show that $\lim_{h \rightarrow \infty} \pi(h) = 0$, implying $\lim_{h \rightarrow \infty} \chi(h) = 0$.
In Figure \ref{fig:chi}, we give $\chi(h)$ for $K = 1, 3, 5, 10$ partitions for a skew-$t$ distribution with $\alpha = 10$, and 3 degrees of freedom.
In Figure \ref{fig:chi} we estimate $\pi(h)$ through simulation.
% To estimate $\pi(h)$, we generate 500 sites uniformly over the unit-square.
% We then randomly generate 400 different sets of partitions using $K = 3$, $5$, and $10$.
% For each set of knots, we take $\pi(h)$ to be the proportion of sites in the same partition that are separated by distance $h$.
% This plot demonstrates how partitioning helps to reduce extremal dependence as $h$ increases.

@@

\subsection{Extension to space-time data} \label{s:temporal}
When using daily measurements, the assumption of temporal independence is often inappropriate.
In this section, we extend (\ref{eq:partition}) to the spatiotemporal setting.
There are several places where temporal dependence could be incorporated in the model, including the residuals $v_t(\bs)$.
However, we choose to allow for temporal dependence in the $\bw$, $z$, and $\sigma$ terms because these terms dictate the tail behavior which is our primary focus.
Let
\begin{align} \label{eq:spatiotemp}
  Y_t(\bs) = \bX_t(\bs)^T \bbeta + \lambda \sigma_t(\bs) |z_t(\bs)| + \sigma_t(\bs) v_t(\bs),
\end{align}
where $t \in \{1, \ldots, T\}$ denotes the day of each observation.
Let \hbox{$\bw_{tk} = (w_{tk1}, w_{tk2})$} be a spatial knot on day $t$, and let $w_{t1}, \ldots, w_{tK}$ be a collection of spatial knots on day $t$.
As in Section \ref{s:part}, these knots define a daily partition $P_{t1}, \ldots, P_{tK}$, and for $\bs \in P_{tk}$,
\begin{align}
  z_t(\bs) = z_{tk}\quad \text{and} \quad \sigma_{t}(\bs) = \sigma_{tk}.
\end{align}
We allow the partition structure to vary from day to day in order to account for sharp spikes in a response that may not be present every day (e.g. the impact of a forest fire on ozone levels).

We use an AR(1) time series model for $w_{tk}$, $z_{tk}$, and $\sigma_{tk}$.
The time series model must be specified after a transformation to preserve the skew-$t$ process at each time point.
For each time-varying parameter, we transform to obtain a standard normal marginal distribution, place a Gaussian prior with autocorrelation on the transformed parameter, and then transform back to the appropriate marginal distribution for the skew-$t$ process.
We first transform the spatial knots from $\calD$ to $\calR^2$ as follows.
Let
\begin{align}
  w^*_{tki} = \Phi^{-1}\left[ \frac{ w_{tki} - \min(\bs_i)}{ \max(\bs_i) - \min(\bs_i) } \right], \quad i = 1, 2
\end{align}
where $\Phi$ is a univariate standard normal density function and $\bs_i = [s_{1i}, \ldots, s_{ni}]$.
Then the transformed knots $\bw^*_{tk} \in \calR^2$.
We use a copula on $\sigma^2_t(\bs)$ to ensure that the marginal distributions of $\sigma^2_t(\bs)$ are inverse gamma.
Let
\begin{align}
  \sigma^{2*}_t(\bs) =\Phi^{-1}\left\{ \text{IG}[\sigma^2_t(\bs)] \right\}
\end{align}
where IG is defined as before.
We also use a copula on $z_{t}(\bs)$ to ensure that the marginal distributions of $z_t(\bs)$ are half-normal.
Let
\begin{align}
  z^*_t(\bs) = \Phi^{-1}\left\{ \text{HN}[z_t(\bs)] \right\}
\end{align}
where HN is the distribution function of a half-normal random variable.
The AR(1) process for each tail parameter is $\bw^*_{1k} \sim N_w(0, 1)$, $z^*_{1k} \sim N(0, \sigma^2_{1k})$, $\sigma^{2*}_{1k} \sim N(0, 1)$, and for $t > 1$ the time series is modeled as
\begin{align}
  \bw^*_{tk} | \bw^*_{t-1, k} &\sim N_2\left[\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2) \right] \\
  z^*_{tk} | z^*_{t-1, k} &\sim N \left[\phi_z z^*_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2)\right] \\
  \sigma^{2*}_{tk} | \sigma^{2*}_{t-1, k} &\sim N \left[\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2) \right]
\end{align}
where $|\phi_w|$, $|\phi_z|$, $|\phi_\sigma| < 1$.
These are stationary time series models with marginal distributions \hbox{$\bw^*_{k} \sim N_2(0, 1)$}, \hbox{$z^*_{k} \sim N(0, \sigma^2_{k})$}, and \hbox{$\sigma^{2*}_{k} \sim N(0, 1)$}.
After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $z_{tk} \sim HN(0, \sigma^2_{tk})$ $\sigma^2_{tk} \sim \text{IG}(a, b)$.
For each day, the model is identical to the spatial-only model in (\ref{eq:partition}) by construction.

\section{Hierarchical model}\label{s:hier}
Conditioned on $z_{tk}(\bs)$, $\sigma^2_{tk}(\bs)$, and $P_{tk}$, the marginal distributions are Gaussian and the joint distribution multivariate Gaussian.
However, we do not fix the partitions, they are treated as unknown and updated in the MCMC algorithm.
We model this with a Bayesian hierarchical model as follows.
Let $\bw_{t1}, \ldots, \bw_{tK}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, and $P_{tk}$ as defined in (\ref{eq:subregions}).
In practice, we fix $K$ at different levels, and assess its impact on prediction as described in \ref{s:modelselect}.
Then
\begin{align}
   Y_t(\bs) \mid z_{t}(\bs), \sigma_t^2(\bs), P_{tk}, \Theta &= \bX_t(\bs)^T \beta + \lambda |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
   z_t(\bs) &= z_{tk} \text{ if } \bs \in P_{tk}\nonumber\\
   \sigma^2_{t}(\bs) &= \sigma^2_{tk} \text{ if } \bs \in P_{tk}\nonumber\\
   \lambda &= \lambda_1 \lambda_2\nonumber\\
   \lambda_1 &= \left\{ \begin{array}{ll}
      +1 \quad & \text{w.p. } 0.5\\
      -1 \quad & \text{w.p. } 0.5
   \end{array}\right.\nonumber\\
   \lambda^2_2 & \sim IG(a, b)\nonumber\\
   v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\nonumber\\
   z^*_{tk} \mid z^*_{t-1, k}, \sigma^2_{tk} &\sim N(\phi_z z^*_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2))\nonumber\\
   \sigma^{2*}_{tk} \mid \sigma^{2*}_{t-1, k} &\sim N(\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2))\nonumber\\
   \bw^*_{tk} \mid \bw^*_{t-1, k} &\sim N_2(\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2)) \nonumber
\end{align}
where $\Theta = \{\rho, \nu, \gamma\, \lambda, \beta\}$, and $\Sigma$ is a \Matern covariance matrix as described in Section \ref{s:skewt}.
We parameterize $\lambda = \lambda_1 \lambda_2$ to help with convergence in the MCMC.

>>>>

% cross-referencing appendices
\usepackage{xr}
\externaldocument{supplemental-rev}

@@

Our model allows for inference and predictions using the full likelihood with computing on the order of Gaussian models.
This allows us to use Bayesian methods to impute data below the threshold as well as to make predictions at unobserved locations.
In a spatial setting, the multivariate skew-$t$ distribution demonstrates asymptotic dependence between observations at all sites regardless of the distance between the sites.
In order to address this concern, we introduce a random spatial partition similar to the method used by \citet{Kim2005} for non-stationary Gaussian data.
This partitioning alleviates the asymptotic spatial dependence present in the skew-$t$ distribution for sites that are far apart.

The paper is organized as follows.
Section \ref{s:spatialskew} is a brief review of the spatial skew-$t$ process.
In Section \ref{s:spatial}, we build upon the traditional skew-$t$ process by incorporating censoring to focus on tails, partitioning to remove long-range asymptotic dependence, and extending the model to space-time data.
The computing is described in Section \ref{s:hier}.
In Section \ref{s:simstudy}, we present a simulation study that examines the predictive capabilities of this model compared to Gaussian and max-stable methods.
We then compare our method to Gaussian and max-stable methods with a data analysis of ozone measurements from 800 sites throughout the US in Section \ref{s:analysis}.
The final section provides brief discussion and direction for future research.

@@

\subsection{Skew-$t$ process} \label{s:skewt}
% Let $Y(\bs)$ be the observation at spatial location $\bs$ in a spatial domain of interest $\calD \in \calR^2$.
% The spatial skew-$t$ process can be written
% \begin{align}
%   Y(\bs) = \bX(\bs)^T \bbeta + \lambda \sigma |z| + \sigma v(\bs)
% \end{align}
% where $\bX(\bs)$ is a set of spatial covariates at site $\bs$, $\bbeta$ is the vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a, b)$ is random scale parameter, IG is the distribution function of an inverse gamma random variable, and $v(\bs)$ is a spatial Gaussian process with mean zero, variance one, and a positive definite correlation function.

% For a finite collection of locations $\bs_1, \ldots, \bs_n$, denote the vector of observations $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^T$.
% After marginalizing over both $z$ and $\sigma$,
% \begin{align}
%   \bY \sim \text{ST}_n(\bX \bbeta, \bOmega, \balpha, 2a), \label{eq:fullmodel}
% \end{align}
% that is, $\bY$ follows an $n$-dimensional skew-$t$ distribution with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$, and $\bSigma$ is the positive definite correlation matrix of $[v(\bs_1), \ldots, v(\bs_n)]$.
% This process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $2a$.
% Furthermore, the marginal distributions at each location also follow a univariate skew-$t$ distribution \citep{Azzalini2014}.
Various forms of multivariate skew-normal and skew-$t$ distributions have been proposed in the literature.
In this section we introduce a popular version \citep{Azzalini2014,Beranger2016}, and demonstrate how the connection between this parameterization and one that we propose to use.
\citet{Azzalini2014} and \citet{Beranger2016} define a skew-normal process as
\begin{align}\label{eq:origskewnorm}
  \tilde{X}(\bs) = \tilde{\lambda} |z| + (1 - \tilde{\lambda}^2)^{1 / 2} v(\bs), \quad \bs \in \calD
\end{align}
where $\tilde{\lambda} \in (-1, 1)$, $z \sim N(0, 1)$, $v(\bs)$ is a Gaussian process with mean zero, variance one, and spatial correlation function cor$(h)$ where $h = || \bs_1 - \bs_2 || \in \calR$ is the distance between sites $\bs_1$ and $\bs_2$, and $\calD \in \calR^2$ is the spatial domain of interest.
To extend the skew-normal process to a skew-$t$ process \citet{Azzalini2003} take
\begin{align} \label{eq:originalskewt}
  \tilde{Y}(\bs) = \xi + V X(\bs)
\end{align}
where $V^{-2} \sim$ Gamma$(\nu / 2, \nu / 2)$.
Now, let $\xi = \bX(\bs)^\prime \bbeta$ where $\bX(\bs)^\prime$ is a set of $p$ spatial covariates at site $\bs$, $\bbeta$ is the $p$-vector of regression parameters, $\sigma^{-2} = \frac{a}{b} V^{-2}$ where $a = \nu / 2$ and $b = \frac{\nu}{2 (1 - \tilde{\lambda}^2)} > 0$, and let $\lambda = \tilde{\lambda} / \sqrt{1 - \tilde{\lambda}^2} \in \calR$.
Then we can express (\ref{eq:originalskewt}) as
\begin{align}
  Y(\bs) = \bX(\bs)^\prime \bbeta + \lambda \sigma |z| + \sigma v(\bs)
\end{align}
where $\sigma^2 \sim $IG$(a, b)$.

Although any positive definite correlation function could be used, we choose to use the stationary isotropic \Matern correlation with
\begin{align}
  \text{cor}(h) = (1 - \gamma) I(h = 0) + \gamma \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right) \label{eq:matern}
\end{align}
where $\rho$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs_1 - \bs_2 ||$ is the distance between sites $\bs_1$ and $\bs_2$.
We use this parameterization for spatial correlation because the $\gamma$ parameter permits the inclusion of a nugget effect to account for non-spatial variability due to issues like measurement error.

\subsection{Extremal dependence}\label{s:extdep}
Our interest lies in spatial dependence in the tail of the skew-$t$ process.
One measure of extremal dependence is the $\chi$ statistic \citep{Coles1999}.
For a stationary and isotropic spatial process, the $\chi$ statistic for locations $\bs_1$ and $\bs_2$ separated by distance $h = || \bs_1 - \bs_2 ||$ with identical marginal distributions is
\begin{align}
  \chi(h) = \lim_{c \rightarrow c^*} \Pr[Y(\bs_1) > c | Y(\bs_2) > c]
\end{align}
where $c^*$ is the upper limit of the support of $Y$.
For the skew-$t$ distribution, $c^* = \infty$.
If $\chi(h) = 0$, then observations are asymptotically independent at distance $h$.
For Gaussian processes, $\chi(h) = 0$ regardless of the distance $h$, so they are not suitable for modeling asymptotically dependent extremes.
Unlike the Gaussian process, the skew-$t$ process is asymptotically dependent (the explicit expression for $\chi(h)$ is given in \ref{a:skewt}).
However, one problem with the spatial skew-$t$ process is that $\lim_{h \rightarrow \infty} \chi(h) > 0$.
This occurs because all observations, both near and far, share the same $z$ and $\sigma$ terms.
Therefore, this long-range dependence feature of the skew-$t$ process is not ideal for spatial analysis of large geographic regions where we expect only local spatial dependence.
We propose a solution to this in Section \ref{s:part}.

@@

\subsection{Partitioning to remove long-range asymptotic dependence}\label{s:part}
The motivation for the partition is that for a large spatial domain, it may not be reasonable to assume sites that are far apart demonstrate asymptotic dependence.
Modeling different levels of asymptotic dependence was discussed by \citet{Wadsworth2012} with a hybrid spatial dependence model.
\citet{Huser2014} also allow for varying asymptotic dependence across both space and time with a partition structure represented by random discs moving across the space for a random duration with a random velocity and random radius.
We handle the problem of long-range asymptotic dependence with a random partition.
As discussed in Section \ref{s:spatialskew}, the source of long-range dependence is the shared $z$ and $\sigma$.
Therefore, to alleviate this dependence, we allow $z$ and $\sigma$ to vary by site.
The model becomes
\begin{align} \label{eq:partition}
  Y(\bs) &= \bX(\bs)^T \bbeta + \lambda \sigma(\bs) |z(\bs)| + \sigma(\bs) v(\bs).
\end{align}
To model spatial variation, consider a set of spatial knots $\bw_1, \ldots, \bw_K$ from a homogeneous Poisson process with intensity $\mu$ over spatial domain $\calD \in \calR^2$.
The knots define a random partition of $\calD$ by subregions $P_{1}, \ldots, P_{K}$ defined as
\begin{align} \label{eq:subregions}
  P_{k} = \{ \bs : k = \argmin_\ell || \bs - \bw_{\ell} || \}.
\end{align}
In other words, $P_1$ is composed of all sites for which the closest knot is $\bw_1$.
All $z(\bs)$ and $\sigma(\bs)$ for sites in subregion $k$ are assigned common values
\begin{align}
  z(\bs) = z_{k} \quad \text{and} \quad \sigma(\bs) = \sigma_{k} \label{eq:sitezsig}
\end{align}
and the $z_k$ and $\sigma^2_k$ are distributed as $z_k \iid N(0, 1)$ and $\sigma^2_k \iid$ IG$(a, b)$.
So, within each partition, $Y(\bs)$ follows the spatial skew-$t$ process defined in Section \ref{s:spatialskew}.
Across partitions, the $Y(\bs)$ remain correlated via the correlation function for $v(\bs)$ because $v(\bs)$ spans all partitions.

The partitioning model remove long-range dependence. Conditional on knots $\bw_1, \ldots, \bw_K$, the $\chi$ statistic for two sites $\bs_1$ and $\bs_2$ in partitions $k_1$ and $k_2$ respectively is
\begin{align}
  \chi(h) &= I(k_1 = k_2) \chi_{\text{skew-}t}(h) + I(k_1 \neq k_2) \chi_{\text{Gaus}}(h) \nonumber \\
         &= I(k_1 = k_2) \chi_{\text{skew-}t}(h)
\end{align}
where $I(\cdot)$ is an indicator function, $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process given in (\ref{eq:chiskew-t}), $\chi_{\text{Gaus}}(h)$ is the $\chi$ statistic for a Gaussian process, and $h = ||\bs - \bt||$.
Therefore, sites in different subregions are asymptotically independent because $\chi_{\text{Gaus}}(h) = 0$ for all $h$.
Marginally, over the knots, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_1 = k_2)$ is the probability that two sites separated by distance $h$ are in the same partition.
In \ref{a:proofsamepartition}, we show that $\lim_{h \rightarrow \infty} \pi(h) = 0$, implying $\lim_{h \rightarrow \infty} \chi(h) = 0$.
In Figure \ref{fig:chi}, we give $\chi(h)$ for $K = 1, 3, 5, 10$ partitions for a skew-$t$ distribution with $\alpha = 10$, and 3 degrees of freedom.
In Figure \ref{fig:chi} we estimate $\pi(h)$ through simulation.
% To estimate $\pi(h)$, we generate 500 sites uniformly over the unit-square.
% We then randomly generate 400 different sets of partitions using $K = 3$, $5$, and $10$.
% For each set of knots, we take $\pi(h)$ to be the proportion of sites in the same partition that are separated by distance $h$.
% This plot demonstrates how partitioning helps to reduce extremal dependence as $h$ increases.

@@

\subsection{Extension to space-time data} \label{s:temporal}
When using daily measurements, the assumption of temporal independence is often inappropriate.
In this section, we extend (\ref{eq:partition}) to the spatiotemporal setting.
There are several places where temporal dependence could be incorporated in the model, including the residuals $v_t(\bs)$.
However, we choose to allow for temporal dependence in the $\bw$, $z$, and $\sigma$ terms because these terms dictate the tail behavior which is our primary focus.
Let
\begin{align} \label{eq:spatiotemp}
  Y_t(\bs) = \bX_t(\bs)^T \bbeta + \lambda \sigma_t(\bs) |z_t(\bs)| + \sigma_t(\bs) v_t(\bs),
\end{align}
where $t \in \{1, \ldots, T\}$ denotes the day of each observation.
Let \hbox{$\bw_{tk} = (w_{tk1}, w_{tk2})$} be a spatial knot on day $t$, and let $\bw_{t1}, \ldots, \bw_{tK}$ be a collection of spatial knots on day $t$.
As in Section \ref{s:part}, these knots define a daily partition $P_{t1}, \ldots, P_{tK}$, and for $\bs \in P_{tk}$,
\begin{align}
  z_t(\bs) = z_{tk}\quad \text{and} \quad \sigma_{t}(\bs) = \sigma_{tk}.
\end{align}
We allow the partition structure to vary from day to day in order to account for sharp spikes in a response that may not be present every day (e.g. the impact of a forest fire on ozone levels).

We use an AR(1) time series model for $\bw_{tk}$, $z_{tk}$, and $\sigma_{tk}$.
The time series model must be specified after a transformation to preserve the skew-$t$ process at each time point.
For each time-varying parameter, we transform to obtain a standard normal marginal distribution, place a Gaussian prior with autocorrelation on the transformed parameter, and then transform back to the appropriate marginal distribution for the skew-$t$ process.
We first transform the spatial knots from $\calD$ to $\calR^2$ as follows.
Let
\begin{align}
  w^*_{tki} = \Phi^{-1}\left[ \frac{ w_{tki} - \min(\bs_i)}{ \max(\bs_i) - \min(\bs_i) } \right], \quad i = 1, 2
\end{align}
where $\Phi$ is a univariate standard normal density function and $\bs_i = [s_{1i}, \ldots, s_{ni}]$.
Then the transformed knots $\bw^*_{tk} \in \calR^2$.
We use a transformation of a Gaussian random variable on $\sigma^2_t(\bs)$ to ensure that the marginal distributions of $\sigma^2_t(\bs)$ are inverse gamma.
Let
\begin{align}
  \sigma^{2*}_t(\bs) =\Phi^{-1}\left\{ \text{IG}[\sigma^2_t(\bs)] \right\}
\end{align}
where IG is defined as before.
We also use a transformation of a Gaussian random variable on $z_{t}(\bs)$ to ensure that the marginal distributions of $z_t(\bs)$ are half-normal.
Let
\begin{align}
  z^*_t(\bs) = \Phi^{-1}\left\{ \text{HN}[z_t(\bs)] \right\}
\end{align}
where HN is the distribution function of a half-normal random variable.
The AR(1) process for each tail parameter is $\bw^*_{1k} \sim N_w(0, 1)$, $z^*_{1k} \sim N(0, 1)$, $\sigma^{2*}_{1k} \sim N(0, 1)$, and for $t > 1$ the time series is modeled as
\begin{align}
  \bw^*_{tk} | \bw^*_{t-1, k} &\sim N_2\left[\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2) \right] \\
  z^*_{tk} | z^*_{t-1, k} &\sim N \left[\phi_z z^*_{t-1, k}, (1 - \phi_z^2)\right] \\
  \sigma^{2*}_{tk} | \sigma^{2*}_{t-1, k} &\sim N \left[\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2) \right]
\end{align}
where $|\phi_w|$, $|\phi_z|$, $|\phi_\sigma| < 1$.
These are stationary time series models with marginal distributions \hbox{$\bw^*_{k} \sim N_2(0, 1)$}, \hbox{$z^*_{k} \sim N(0, 1)$}, and \hbox{$\sigma^{2*}_{k} \sim N(0, 1)$}.
After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $z_{tk} \sim HN(0, 1)$. and $\sigma^2_{tk} \sim \text{IG}(a, b)$.
For each day, the model is identical to the spatial-only model in (\ref{eq:partition}) by construction.

\section{Hierarchical model}\label{s:hier}
Conditioned on $z_{tk}(\bs)$, $\sigma^2_{tk}(\bs)$, and $P_{tk}$, the marginal distributions of $Y(\bs)$ are Gaussian and the joint distribution of $Y(\bs)$ is multivariate Gaussian.
However, we do not fix the partitions, they are treated as unknown and updated in the MCMC algorithm.
We model this with a Bayesian hierarchical model as follows.
Let $\bw_{t1}, \ldots, \bw_{tK}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, and $P_{tk}$ as defined in (\ref{eq:subregions}).
In practice, we fix $K$ at different levels, and assess its impact on prediction as described in \ref{s:modelselect}.
Then
\begin{align}
   Y_t(\bs) \mid z_{t}(\bs), \sigma_t^2(\bs), P_{tk}, \Theta &= \bX_t(\bs)^T \beta + \lambda \sigma_t(\bs) |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
   z_t(\bs) &= z_{tk} \text{ if } \bs \in P_{tk}\nonumber\\
   \sigma^2_{t}(\bs) &= \sigma^2_{tk} \text{ if } \bs \in P_{tk}\nonumber\\
   \lambda & \sim N(0, 20)\nonumber\\
   % \lambda_1 &= \left\{ \begin{array}{ll}
   %    +1 \quad & \text{w.p. } 0.5\\
   %    -1 \quad & \text{w.p. } 0.5
   % \end{array}\right.\nonumber\\
   % \lambda^2_2 & \sim IG(a, b)\nonumber\\
   v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\nonumber\\
   z^*_{tk} \mid z^*_{t-1, k} &\sim N(\phi_z z^*_{t-1, k}, (1 - \phi_z^2))\nonumber\\
   \sigma^{2*}_{tk} \mid \sigma^{2*}_{t-1, k} &\sim N(\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2))\nonumber\\
   \bw^*_{tk} \mid \bw^*_{t-1, k} &\sim N_2(\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2)) \nonumber
\end{align}
where $\Theta = \{\rho, \nu, \gamma\, \lambda, \beta\}$, and $\Sigma$ is a \Matern covariance matrix as described in Section \ref{s:skewt}.
